Consistent Polyhedral Surrogates via Embedding

### 2019-10-23 notes/top-k.tex
Minor changes, fixing on of the align proofs

We write the expected loss $\inprod{p}{L^{top-k}(u)}$ in a linear form without the $(\cdot)_+$ operator.
Then, using this linear form of the expected loss, we are able to find a finite set of optimal reports over all $p \in \simplex$.

@@

	Now suppose that $y < k+1$ and $u_y \in [0,1)$; we now observe $(u - e_y)_{[k]} = u_{[k+1]} = 0$. 
	\begin{align*}
	L^{top-k}(u)_y &= \left(1 - u_y + \frac 1 k \sum_{i=1}^{k-1} u_{[i]} + 0 \right)_+\\
	&= \left(1 - u'_y + \frac 1 k \sum_{i=1}^{k-1} u'_{[i]}\right)_+\\
	&= \left(1 - u'_y + \frac 1 k \sum_{i=1}^{k} (u' - e_y)_{[i]}\right)_+\\
	&= L^{top-k}(u')_y~.~
	\end{align*}

@@

To see this, consider $u_y = \bar u_{-y} + 1 + \epsilon$ and $u'_y = \bar u_{-y} + 1$, with $u_i = u'_i \; \forall i \neq y$. 
\begin{align*}
L^{top-k}(u)_y &= \left(1 - (\bar u_{-y} + \epsilon + 1) + \frac 1 k \left(\sum_{i=1, i \neq y}^k u_i + \bar u_{-y} + \epsilon\right) \right)_+\\ 
&= \left(1 - (\bar u_{-y} + \epsilon + 1) + \frac 1 k \left((k-1) \bar u_{-y} + \bar u_{-y} + \epsilon\right) \right)_+\\
&= \left(-\bar u_{-y} - \epsilon + \bar u_{-y} + \frac \epsilon k \right)_+ \\
&= (-\frac{k-1}{k}\epsilon)_+\\
&= 0
\end{align*}
Compare the above to the following:
\begin{align*}
L^{top-k}(u')_y &= \left(1 - (\bar u_{-y} + 1) + \frac 1 k \left(\sum_{i=1, i \neq y}^k u_i + \bar u_{-y} \right) \right)_+\\ 
&= \left(1 - (\bar u_{-y} + 1) + \frac 1 k \left((k-1) \bar u_{-y} + \bar u_{-y} \right) \right)_+\\
&= \left(-\bar u_{-y} + \bar u_{-y}\right)_+ \\
&= 0~.~
\end{align*}
Thus, the losses are equal on the outcome $y$ so that $u_y > \bar u_{-y} + 1$.

Now let us consider outcome $j \neq y$.
By Assumption~\ref{assum:nonneg}, if $u_y > \bar u_{-y} + 1$, then we have $y < k+1$ as $u_y > 0$.
Now, for outcome $j \neq y$, we have 
\begin{align*}
L^{top-k}(u)_j &= \left( 1 - u_j + \frac 1 k \sum_{i=1}^k (u-e_j)_{[i]} \right)_+\\
	&= \left( 1 - u'_j + \frac 1 k \sum_{i=1}^k (u'-e_j)_{[i]} + \frac {\epsilon}{k}\right)_+\\
	&> \left( 1 - u'_j + \frac 1 k \sum_{i=1}^k (u'-e_j)_{[i]}\right)_+\\
	&= L^{top-k}(u')_j
\end{align*}
Therefore, if $u$ minimizes $L^{top-k}(\cdot)_y$, then so does $u'$.
\end{proof}

@@

\section{Re-writing the loss}

@@

	If $y \geq k+1$, then $\frac 1 k \sum_{i=1}^k (u-e_y)_{[i]} = \bar u$, so the inequality becomes $u_y \leq 1 + \bar u$.
	This holds since $u_y \leq u_k \leq \bar u \leq 1 + \bar u$ by Assumptions~\ref{assum:ordered-u} and~\ref{assum:nonneg}.

@@

	Now if $u_y > 1$, then $\frac 1 k \sum_{i=1}^k (u-e_y)_{[i]} = \bar u - \frac 1 k$.
	Therefore, we have 
	\begin{align*}
	&1 -u_y + \bar u - \frac 1 k \geq 0\\
	&\iff 1 + \frac{k}{k-1}\bar u_{-y} \geq u_y + \frac 1 k\\
	&\iff \frac{1}{k-1} \geq \frac 1 k~,~
	\end{align*}
	with the final if and only if statement coming from Assumption~\ref{assum:top-threshold}.

@@

\begin{theorem}\label{thm:finite-set}
	For the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}, for all $p \in \simplex$, there is a report in the following report set $\R$ minimizing $\inprod{p}{L^{top-k}(\cdot)}$.
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
	\jessie{Check the $|H|+|M| \leq k$ condition. Worst case, the set $\R$ is still finite without it.}
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by \jessie{???}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.

	To see the reason for $c := \frac{|M| + k -1}{k - |H|}$, consider that, given the loss $L^{top-k}$, when we want to ``crank $u_i$'' above $1$, we want to do so as much as possible, until hitting the threshold discussed in Assumption~\ref{assum:top-threshold}.
	Exceeding the threshold on $u_i > c$ then can only increase expected loss.

>>>>

We write the expected loss $L^{top-k}(u)$ without the $(\cdot)_+$ operator by restricting $u$.
Then, using this form of the expected loss, we are able to find a finite set of optimal reports over all $p \in \simplex$.

@@

	Now suppose that $y < k+1$ and $u_y \in [0,1)$; we now observe $(u - e_y)_{[k]} = u_{k+1} = 0$. 
	\begin{align*}
	L^{top-k}(u)_y &= \left(1 - u_y + \frac 1 k \sum_{i=1}^{k-1} u_{[i]} + 0 \right)_+\\
	&= \left(1 - u'_y + \frac 1 k \sum_{i=1}^{k-1} u'_{[i]}\right)_+\\
	&= \left(1 - u'_y + \frac 1 k \sum_{i=1}^{k} (u' - e_y)_{[i]}\right)_+\\
	&= L^{top-k}(u')_y~.~
	\end{align*}

@@

To see this, consider $u_y = \bar u_{-y} + 1 + \epsilon$ and $u'_y = \bar u_{-y} + 1$, with $u_i = u'_i$ for all $i \neq y$. 
\begin{align*}
L^{top-k}(u)_y &= \left(1 - (\bar u_{-y} + \epsilon + 1) + \frac 1 k \left(\sum_{i=1, i \neq y}^k u_i + \bar u_{-y} + \epsilon\right) \right)_+\\ 
&= \left(1 - (\bar u_{-y} + \epsilon + 1) + \frac 1 k \left((k-1) \bar u_{-y} + \bar u_{-y} + \epsilon\right) \right)_+\\
&= \left(-\bar u_{-y} - \epsilon + \bar u_{-y} + \frac \epsilon k \right)_+ \\
&= (-\frac{k-1}{k}\epsilon)_+\\
&= 0
\end{align*}
Compare the above to the following:
\begin{align*}
L^{top-k}(u')_y &= \left(1 - (\bar u_{-y} + 1) + \frac 1 k \left(\sum_{i=1, i \neq y}^k u_i + \bar u_{-y} \right) \right)_+\\ 
&= \left(1 - (\bar u_{-y} + 1) + \frac 1 k \left((k-1) \bar u_{-y} + \bar u_{-y} \right) \right)_+\\
&= \left(-\bar u_{-y} + \bar u_{-y}\right)_+ \\
&= 0~.~
\end{align*}
Thus, the losses are equal on the outcome $y$ where $u_y > \bar u_{-y} + 1$.

Now let us consider outcome $j \neq y$.
By Assumption~\ref{assum:nonneg}, if $u_y > \bar u_{-y} + 1$, then we have $y < k+1$ as $u_y > 0$.
Now, for outcome $j \neq y$, we have 
\begin{align*}
L^{top-k}(u)_j &= \left( 1 - u_j + \frac 1 k \sum_{i=1}^k (u-e_j)_{[i]} \right)_+\\
	&= \left( 1 - u'_j + \frac 1 k \sum_{i=1}^k (u'-e_j)_{[i]} + \frac {\epsilon}{k}\right)_+\\
	&> \left( 1 - u'_j + \frac 1 k \sum_{i=1}^k (u'-e_j)_{[i]}\right)_+\\
	&= L^{top-k}(u')_j
\end{align*}
Therefore, if $u$ minimizes $L^{top-k}(\cdot)_y$, then so does $u'$ for each $y \in \Y$, so we can say the same of the expected loss $\inprod{p}{L^{top-k}(\cdot)}$ for all $p \in \simplex$.
\end{proof}

@@

\section{Re-writing the loss}
The following Lemma allows us to consider the loss without the $(\cdot)_+$ operator.

@@

	If $y \geq k+1$, then $\frac 1 k \sum_{i=1}^k (u-e_y)_{[i]} = \bar u$, so the inequality becomes $u_y \leq 1 + \bar u$.
	This holds since $u_y \leq u_k \leq \bar u \leq 1 + \bar u$ by Assumption~\ref{assum:ordered-u}.

@@

	Now if $u_y > 1$, then $\frac 1 k \sum_{i=1}^k (u-e_y)_{[i]} = \bar u - \frac 1 k$.
	Therefore, we have 
	\begin{align*}
	&1 -u_y + \bar u - \frac 1 k \geq 0\\
	&\iff 1 + \frac{k}{k-1}\bar u_{-y} \geq \frac{k-1}{k} u_y + \frac 1 k\\
	&\iff (1 + \bar u_{-y}) + \frac 1 {k-1} \bar u_{-y} \geq \frac{k-1}{k} u_y + \frac 1 k\\
	&\iff \frac{k-1}{k}(1 + \bar u_{-y}) + \frac 1 k + \frac{1}{k}\bar u_{-y} + \frac 1 {k-1} \bar u_{-y} \geq \frac{k-1}{k} u_y + \frac 1 k\\
	&\implies \frac{1}{k}\bar u_{-y}+ \frac{1}{k-1}\bar u_{-y} \geq 0~,~
	\end{align*}
	with the final implication coming from Assumption~\ref{assum:top-threshold}.

@@

\begin{theorem}\label{thm:finite-set}
	Consider the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}.  
	For all $p \in \simplex$, there is a report in the following report set $\R$ minimizing $\inprod{p}{L^{top-k}(\cdot)}$.
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
	\jessie{Check the $|H|+|M| \leq k$ condition. Worst case, the set $\R$ is still finite without it.}
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by \jessie{???}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.

	To see the reason for $c := \frac{|M| + k -1}{k - |H|}$, consider that, given the loss $L^{top-k}$, when we want to ``crank $u_i$'' above $1$, we want to do so as much as possible, until hitting the threshold discussed in Assumption~\ref{assum:top-threshold}.
	Exceeding the threshold on $u_i > c$ then can only increase expected loss, as shown in Lemma~\ref{lem:top-threshold}.
